Haploid Selection, Sex Ratio Bias, and Transitions Between Sex-Determination Systems

### 2018-05-18 LaTeX/response.tex

\noindent\subsection{}
Likewise, the results yield three major insights:\\
(1) SD transitions can, counterintuitively, potentially INCREASE sex ratio bias\\
(2) Tight linkage and sexual antagonism are not required for transitions to occur\\
(3) Under haploid selection, conditions for the spread of new sex-determining alleles are extremely permissive.\\
Again, all of these points are there, but they could be better drawn out than they currently are. The results, as written, are hard-going, and I expect will be particularly challenging for readers that are not heavily immersed in evolutionary theory.

\noindent\paragraph{Response}
\textcolor{blue}{
As discussed further below (response XXXXX), we have highlighted conclusions in the results section as suggested. 
To further reduce the technicality of our presentation, we have also revised the results section. 
In particular, we have restructured the presentation of the subsections ``Generic invasion by a neo-Y or neo-W'' and ``Tight linkage with the ancestral sex-determining locus''. 
In the ``Generic invasion by a neo-Y or neo-W" section, we have moved some details to the supplementary material (particularly, proofs regarding the leading eigenvalue, formally LXXX-XXX). 
Now, we focus on explaining the intuition behind our technical results.  
For example, we first discuss the `haplotypic growth rates' and 'dissociative forces' (L158-170 and `interpreting this condition...', L173).
We then lay out four specific conclusions (L179-199). 
In the ``Tight linkage'' section, we now first describe specific cases why neo-W's can have improved female fitness, Conclusion 1, L235-259, with more focus on non-technical interpretation (e.g., L237-240). 
Then we discuss the fact that looser sex-linkage can evolve (L260-275, Conclusion 2). 
This re-organisation allows us to break down the logic behind Conclusion 2 more clearly. 
}

@@

\noindent\subsection{}
One potential way to help readers -- particularly those interested in what the paper has to say, but are not mathematically inclined -- is to conspicuously highlight each important result and briefly place it into context of earlier theory. But at the same time, the authors have done a great job of drawing out general mathematical results from a relatively complicated model -- their mathematical insights ARE the results and I am not suggesting that they be minimized or shunted into appendices.

One tactic for including analytical details which specialists will want to see, while allowing less mathematically inclined readers to easily find each main result, is to place short and accessible summaries at the close of each section of the Results. Such a tactic has been put to good use in a hugely influential paper by Allen Orr (1998 Evolution), which I expect the authors will be familiar with. Orr presents the major mathematical results of his analysis alongside prominently labelled punchline paragraphs that summarize each of his main results in a concise and easily understandable manner. If the authors choose to take this approach, please be clear to place the take-home messages in context: clearly state whether the result confirms previous theory, or represents a new result or contradiction of an earlier result.

\noindent\paragraph{Response}
\textcolor{blue}{
This is an excellent suggestion and has greatly improved the clarity of presentation in our manuscript. In the results, we have specifically highlighted our main conclusions in the style of Orr 1998. We have chosen slightly different conclusions from the three suggested by the reviewer: 
%We now give 3 main conclusions at the end of the introduction (L69) which echo those mentioned by the reviewer (looser linkage can evolve, equal selection to increase or decrease the sex ratio, ESD does not always invade biased sex ratio; all of which suggest haploid selection makes conditions for transitions more permissive).
\begin{enumerate}
\item[(1)] Selection on loci in or near the non-recombining region around the ancestral sex-determining locus ($r\approx0$) prevents cis-GSD transitions (XY $\leftrightarrow$ XY, ZW $\leftrightarrow$ ZW) but can spur trans-GSD transitions (XY $\leftrightarrow$ ZW). (L236-238)
\item[(2)] With tight linkage between a selected locus and the ancestral sex-determining locus ($r\approx0$), trans-GSD transitions (XY $\leftrightarrow$ ZW) can be favoured by selection \textit{even if they weaken sex-linkage} ($r<R$), potentially shifting sex determination to a different chromosome ($R=1/2$). 
Such transitions can also lead to the maintenance of multifactorial sex-determination systems. (L290-294)
\item[(3A)] New sex-determining alleles (causing cis-GSD transitions, XY $\leftrightarrow$ XY or ZW $\leftrightarrow$ ZW) are favoured if they arise more closely linked with a locus that experiences (haploid and/or diploid) selection than the ancestral-sex-determining locus is ($R<r$). (L326-328)
\item[(3B)] New sex-determining alleles (causing trans-GSD transitions, XY $\leftrightarrow$ ZW) are favoured if they are linked with an ancestrally-autosomal locus that experiences (haploid and/or diploid) selection ($R<1/2$, $r=1/2$). (L333-335)
\item[(3C)] With haploid selection, new sex-determining alleles (causing trans-GSD transitions, XY $\leftrightarrow$ ZW) can spread even if they arise less closely linked with a locus that experiences selection than the ancestral-sex-determining locus is ($r<R$). (L341-343)
\item[(4)] When selection is weak relative to recombination, the presence of haploid selection equally favors the spread of new sex determination systems that reduce sex-ratio bias (benefiting from Fisherian sex ratio selection) or generate a sex-ratio bias (benefiting from associations with selected alleles). (L350-353)
\item[(5)] Transitions from genetic to environmental sex-determination are not straightforwardly predicted by selection to balance the zygotic sex ratio when haploid selection is present.  (L411-413)
%\vspace{0.25cm}\\
\end{enumerate}
These conclusions are also more clearly laid out in the introduction (L48-49, L74-85) and are referenced directly in summarisations in the discussion (L424-432, L543-554). 
}

\textcolor{blue}{Of these conclusions, (3A), (3B), and (3C) have some precendent. (3A) and (3B) are similar to studies by van Doorn and Kirkpatrick (2007,2010) that do not include haploid selection. These comparisons are made immediately above the conclusions (L316-330) and mentioned elsewhere (L55, L77, L489). 
Including haploid selection, Ubeda et al. (2015) also involved examples similar to (3A) and (3B), except that the starting state was ESD. 
Kozielska et al. (2010) also had numerical examples that involved trans-GSD transitions with ($R>r$), as stated in (3C), but focussed only on sex ratio selection. 
We discuss how our results compare to these studies in the introduction (L58-75), discussion (L433-447) and results (L193, L363-369).
Nevertheless, we have highlighted these conclusions, partly for clarity of communication and partly because our results are more general. 
That is, as stated by the reviewer, these results ``converge on previously-identified conditions for the spread of new SD alleles, while also broadening the range of scenarios and conditions that can favour transitions between SD systems''}
%As (1A), (1B), and (2) have some antecedents in the literature, we discuss these immediately below each conclusion. 
%We also discuss, immediately below each conclusion, its novelty.
%These are all new results, although (1A) was shown for the special case of $r=0$, $R=1/2$, and ploidally-antagonistic selection in Kozielska et al 2010, which we now mention immediately below (1A).
}

@@

\noindent\subsubsection{}
Lines 80-81: This primary finding is completely aligned with the earlier views of theory that are summarized above. Are you sure you want to lead with this as the main emphasis of the study? It understates what is new.

\noindent\paragraph{Response}
\textcolor{blue}{
As suggested above, we have re-written to focus on our most novel findings, which are now laid our more specifically, e.g., L71-85 (see also Responses 2.1 and 2.2). 
}

\noindent\subsubsection{}
M locus is dominant to X. How important is this assumption? Is this compatible with any data (e.g., on feminizing or masculinizing factors in Musca domestica)?

\noindent\paragraph{Response}
\textcolor{blue}{
The `epistatic-dominance' between loci is important to the evolutionary dynamics. For example, Veller et al. (2018, Genetics) show that there is a tendency for drift to occur in the direction of the epistatically-dominant sex-determining locus (mentioned on L545-546). Our results also show that the outcome can depend on which sex-determining locus has an epistatically-dominant effect. For example, the end state of the red line in Fig 5 is the starting condition of the other panel. That is, we find that recurrent invasion can occur by a sequence of sex-determining loci that are epistatically-dominant over their predecessors (requiring haploid selection). 
 }

\textcolor{blue}{
Empirical systems with multi-factoral sex-determination (including \textit{Musca domestica}) do indeed indicate that derived sex-determining systems are often `epistatically-dominant'. 
We now mention this where we introduce this aspect of the model (L98-109). 
We also point to the supplementary Mathematica file, which allows arbitrary `epistatic dominance' to be specified. 
There are some numerical examples with reduced `epistatic dominance' in the supplementary material of van Doorn \& Kirkpatrick (2007, Nature), which we also point to here.
Their findings suggest that the selective advantage is reduced (the modifier mutation effectively has a smaller effect) but the direction of evolution is expected to be qualitatively unchanged. 
}

@@

\noindent\subsubsection{}
Paragraph beginning L104: the sex symbols are not ideal. The combined male/female symbol has been used In other contexts to refer to "male and female" rather than representing either/or

\noindent\paragraph{Response}
\textcolor{blue}{This is true, the male/female symbol is commonly used to represent hermaphrodites, we have replaced the male/female symbol with a circle, which is a basis for both the male and female symbols.}

@@

\noindent\paragraph{Response}
\textcolor{blue}{We have made the suggested change.}

\section{Reviewer 3}

This work (henceforth [SOO]) explores the role of both diploid and haploid selection on the evolution of new sex-chromosomes and or transitions between sex-determination systems. For that purpose, the authors develop a population genetics model of three loci with two alleles in each locus. The first locus X is the ancestral sex-determining locus, the second locus A is the locus under diploid and haploid selection. The third locus M is the novel sex-determining locus.

\noindent\subsection{}
I find the article is not clearly written and could benefit from improving the presentation of the main results. 
I also find the article claims that: (a) the modeling work is novel and (b) the findings are unexpected to be only partially justified.
The works of Kozielska 2014 and Ubeda 2015 model the combined effects of both haploid and diploid selection on the evolution of new sex-chromosomes. The work of [SOO] is thus not novel in considering both haploid and diploid selection on the evolution of new sex-chromosomes. It does generalize previous work by considering sex-specific viability selection and get some nice analytical results.

\noindent\paragraph{Response}
%\textcolor{red}{The abstract and summary still seem to be a little over-selling...}
\textcolor{blue}{
We have performed an extensive re-arrangement to improve the clarity of our writing, especially regarding the main results (see Responses 2.1-2.3). 
As suggested by the reviewer (here and below), we also now give more prominence to the related results by Kozielska et al 2010 and Ubeda et al 2015 (e.g., L56-71, L198, L367-373, L440-446). 
In particular, we introduce these studies on Lines 56-71 and more precisely indicate which of our model features and results are new and unexpected in light of these previous results.
}

\textcolor{blue}{
While related, these two studies considered particular cases that we do not consider at length. 
For example, Ubeda et al. (2015) consider ESD-GSD transitions. 
Nevertheless, the transitions in Ubeda et al. (2015) involve increased linkage with a haploid selected locus, which is precedent for our Conclusions 3A and 3B (which are for GSD-GSD transitions in which linkage with a haploid selected locus is increased).
We now mention this in the Introduction (L59-61), Results (L369-373), and Discussion (L432).  
Similarly, Kozielska et al. (2010) involves a special case of GSD transitions during which sex-linkage goes from perfect to none, which is is an example of our Conclusion 2 (and similar to Conclusion 3C, which are conclusions regarding GSD-GSD transitions in which linkage with a haploid selected locus is decreased).
This is mentioned on lines (L63-64, L303, and L368). 
We believe that our new presentation makes the novel and unexpected features of our results much more apparent. 
For example, despite these precedents, we find the combination of Conclusions 3A, 3B, and 3C to be a striking solution that reconciles the results of Kozielska et al. (2010) and Ubeda et al. (2015). 
We elaborate further on what makes our results novel and surprising in the responses below.
}

@@

\noindent\subsubsection{}
The findings in [SOO] claimed to be surprising (``Surprisingly, we find that neither force (sex ratio selection nor associations with genes that have sex-specific effects) dominates the spread of new sex-determining systems alone.") ... [but this result is rather similar to those in Ubeda et al 2015].

\noindent\paragraph{Response}
\textcolor{blue}{
Here we were referring to the result that $\lambda_{W'}^{XY} = \lambda_{Y'}^{ZW}$ with haploid selection, even when sex-ratio biases evolve to become either stronger or weaker (e.g., $\lambda_{Y'}^{ZW}$ vs. $\lambda_{W'}^{XY}$ with $R<r$ and male meiotic drive) or when sex-ratio biases are either present or absent (e.g., $\lambda_{W'}^{XY}$ vs. $\lambda_{Y'}^{ZW}$ with $r<1/2$, $R=1/2$, and male meiotic drive).  
As mentioned by the reviewer, Step 2 in Ubeda et al 2015 shows that associations with a haploid selected locus can favour a new sex-determining allele despite causing biased sex ratios. 
Nevertheless, it was difficult for us to intuit whether the presence/absence of ancestral sex-ratio bias would have a negligible or overwhelming effect on the evolution of new sex-determining alleles. 
Because we start from GSD, our model allows ancestral sex-ratio biases to build up due to the presence of haploid selected loci near to the ancestral sex-determining locus. 
There are reasons to suspect that ancestral-sex-ratio biases would strongly influence the spread of new sex-determining systems.
For example, step 3 of Ubeda et al involves the spread of an autosomal meiotic drive suppressor favoured by Fisherian sex-ratio selection and Kozielska et al 2010 show that ancestral sex-ratio bias can favour new sex determining systems that balance the sex ratio.
We found our result, that these forces are exactly equal (with selection weak relative to recombination), such that there is no difference between the propensity for XY-ZW or ZW-XY transitions, to be surprising.
To clarify our meaning we have rephrased (in Author summary): ``Surprisingly, we find the two forces (selection to equalize the sex-ratio and the benefits of hitch-hiking with driven alleles that distort the sex ratio) will often be equally strong, and thus neither is sufficient to explain the spread of new sex-determining systems in every case." 
We believe that the re-framing of the Introduction/Discussion and the emphasis given to specific Conclusions will help to clarify this point throughout the paper.
}

@@

\noindent\subsubsection{}
\noindent and unexpected (``Even more unexpectedly, we find that, to spread, new sex-determining alleles do not necessarily have to arise in closer linkage with genes that are differentially selected in males and females"), are rather similar to those in Ubeda et al 2015. This is a problem that repeats itself over the paper

\noindent\paragraph{Response}
\textcolor{blue}{
The Ubeda et al. (2015) results do indeed show that differential selection between male and female diploids is not required for new sex-determining alleles to spread; they show that a new genetic sex-determining allele can spread from ancestral ESD when it is linked to a locus that experiences sex-specific meiotic drive. 
In this sentence, we were attempting to stress that new sex determining alleles can spread when they are \textit{less closely linked or even completely unlinked} to loci under differential male/female haploid and/or diploid selection. 
To our knowledge, this has only been shown for the special case of male meiotic drive in an ancestral XY system with $r=0$ and $R=1/2$ with a neo-W, which is driven by sex-ratio selection (Kozielska et al 2010). 
We therefore found the large parameter space allowing invasion with $r<R$, sometimes in the absence of sex-ratio selection (or even in the absence of haploid selection), to be very unexpected.
To improve our communication of this result, we have rephrased (in Author Summary): ``We also find that new sex-determining alleles can spread despite being less closely linked to selected loci, driven by initially tight linkage or haploid selection." 
The changes to the presentation of the Introduction, Results, and Discussion should also help to clarify this point further (see Responses 2.1 and 2.2), this result is now also highlighted as `Conclusion 4'. 
}

@@

\noindent\subsubsection{}
For example in: Page 2. The authors claim that predominant theories of the evolution of sex-determining systems are two: sexually-antagonistic selection and sex-ratio selection. They ignore other two: inbreeding (Charlesworth 1978) and sex-different drive (Ubeda et al 2015).

\noindent\paragraph{Response}
\textcolor{blue}{
Here, sexually-antagonistic selection was referred to as an example of ``selective differences between the sexes". 
We believe the phrase 'selective differences between the sexes' includes sex-differences in drive (Ubeda et al 2015), which occurs during haploid rather than diploid selection, as indicated in Fig 1B (see also Response 3.1.6).
As mentioned elsewhere, the enhanced discussion of Ubeda et al. (2015) should help to clarify this point. 
Charlesworth and Charlesworth 1978 (Am Nat and Heredity) discuss the role of inbreeding (and particularly inbreeding depression) in transitions from hermaphroditism to dioecy. 
As described in the model section and lines 468-472 of the original submission, we have only considered GSD-GSD transitions and GSD-ESD transitions.
For clarity, we have moved this description into the Introduction (L49-50). 
Where we draw a parallel between ESD and hermaphroditism (L468-476), we suggest future theory include inbreeding and cite Charlesworth and Charlesworth (1978). 
}

@@

\noindent\subsubsection{}
Lines 68-69: ``If we assume that haploid selection at any particular locus predominantly occurs in one sex (e.g., meiotic drive during spermatogenesis), then such loci experience a form of sex-specific selection."

This is a stretch and creates confusion. I consider selection acting on individuals with different sex-specific selection, now when selection is acting on gametes (which are neither males or females) males or females are just different environments in which selection may take place. Neither the dynamics nor the constraints are the same.

\noindent\paragraph{Response}
\textcolor{blue}{In this study, we are referring to organisms, such as animals and angiosperms, in which the type of gamete depends on the sex of the diploid that produced it. In this respect, one could consider gametic selection among sperm, for example, to be a form of male-specific selection. 
However, as noted by the reviewer here, the dynamics of selection during gametic selection cannot be captured by a diploid selection coefficient, which is why we specifically model selection upon haploid genotypes.
In our new manuscript, a roughly corresponding statement now reads ``On one hand, haploid selection is typically sex-limited in that it usually occurs among gametes produced by one sex only [38-41].''.
We have also expanded our explanation of the mating system and the way we model haploid selection (see Response 3.2 below).
}

@@

\noindent\subsubsection{}
Lines 70-71: ``In this respect, we might expect that haploid selection would affect transitions between sex-determining systems in a similar manner to sex-specific diploid selection (as explored in [31,32])"

Following up my previous comment Ubeda 2015 explores this and shows that both forms of selection are different. For example sex-antagonistic selection is needed for the invasion of a new sex-determining gene with diploid selection while antagonistic drive in each of the sexes is not needed for the invasion of a new sex-determining gene with haploid selection.

### 2018-05-18 LaTeX/main_revised.tex
improving response letter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Tight linkage allows looser sex-linkage to evolve, even in absence of haploid selection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=1.4\linewidth, right]{../Plots/PositionPlot_SexAntagTighter.eps}}
\caption{
{\bf Transitions between XY and ZW systems can occur even when the new sex-determining locus is less tightly linked to a locus under sexually-antagonistic selection (no haploid selection).}
In panel A, linkage is initially tight relative to selection and a neo-W can invade even when it is less tightly linked with the selected locus ($r<R$; unshaded region around \textbf{*}).
In panel B, linkage is loose enough relative to selection that the analytical results assuming weak selection hold, and a neo-W allele can only invade when it arises at a locus more tightly linked with the selected locus ($R<r$; shaded region).
In panel C we vary the recombination rate between the neo-W and the selected locus ($R$) for a fixed recombination rate between the ancestral sex-determining locus and the selected locus ($r=0.005$). 
Coloured markers show recombination rates for which the temporal dynamics of invasion are plotted in the inset, demonstrating that neo-W alleles reach fixation if they are more (black) or less (red) closely linked to a locus experiencing sexually-antagonistic selection. 
A very loosely linked neo-W does not spread in this case (blue and green lines overlap and go to 0 in inset). 
Fitness parameters are: $w_{AA}^\female = 1.05$, $w_{aa}^\male = 1.2$, $w_{aa}^\female = w_{AA}^\male = 0.85$, $w_{Aa}^\circ = 1$.%,  $t^\circ = \alpha^\circ_\Delta = 0$.
%\textcolor{red}{consider removing panel A, which is repeated in Fig 3.}
%\textcolor{blue}{probably better to remove from panel 3, eh?}
}
\label{fig:SexAntagTighter}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{equation}
\lambda_{Y'}^{(XY)} = 1 + \frac{1}{4}\bar{p}(1-\bar{p})S_{A}^2\frac{ \left( r-R \right) }{r R}+O\left(\epsilon^3 \right) 
\label{eq:lambda_neoY}
\end{equation}

@@

%The new sex-determining allele, $m$, will spread in an XY system if $\lambda_{m}^{(XY)}>1$. 
Eq \eqref{eq:lambda_neoY} demonstrates that, under weak selection, a neo-Y allele will invade an XY system ($\lambda_{Y'}^{(XY)}>1$) if and only if it is more closely linked to the selected locus than the ancestral sex-determining locus (i.e., if $R<r$; note that $\bar{p}(1-\bar{p})S_{A}^2$ is strictly positive as long as $\mathbf{A}$ is polymorphic). 
This echoes our results above where a neo-Y could never invade if $r\approx0$. 
It is also consistent with the results of \cite{vanDoorn:2007eu}, who considered diploid selection only and also found that cis-GSD transitions can only occur when the new sex-determining locus is more closely linked to a locus under sexually-antagonistic selection. 
%Because tighter linkage between the sex-determining region and a selected locus evolves during these transitions, stronger associations between selected alleles and diploid sexes can build up by selection (e.g., between male-beneficial alleles and the male-determining allele, Y).%, increasing diploid mean fitness (Fig \ref{fig:Combination_MeanFit}).

@@

\noindent
Similarly, a neo-W is typically favoured when it is more closely linked to the selected locus than the ancestral sex-determining region is, ($R<r$, e.g., Figs \ref{fig:SexAntagTighter}B and \ref{fig:Combination_Centimorgans}).
Specifically, if the selected locus ($\mathbf{A}$) is ancestrally autosomal ($r=1/2$, leading to $ \hat{p}^\male_Y-\hat{p}^\male_X =0$ such that $\lambda_{W'}^{(XY)}=\lambda_{Y'}^{(XY)}$, Eqs \eqref{eq:lambda_neoY} and Eq \eqref{eq:lambda_neoW}), a more closely linked neo-W ($R<1/2$) is favoured.

@@

\noindent
In the absence of haploid selection ($t^\circ=\alpha^\circ_{\Delta}=0$), the spread of a neo-W is equivalent to the spread of a neo-Y ($\lambda_{W'}^{(XY)}=\lambda_{Y'}^{(XY)}$), such that trans-GSD transitions can occur if and only if the new sex-determining locus is more closely linked to a locus under selection ($R<r$), as found by \cite{vanDoorn:2010hu}. 
However, with haploid selection and some ancestral sex-linkage ($r<1/2$; allowing allele frequency differences on the X and Y), the term in square brackets in Eq \eqref{eq:lambda_neoW} can be positive.
Thus,

@@

To clarify the parameter space under which invasion occurs with weak selection ($\lambda_{W'}^{(XY)}>1$) and $R>r$, we focus on cases where dominance coefficients are equal in the two sexes, $h^\female=h^\male$, and haploid selection only occurs in one sex (e.g., during male meiosis only).
Table \ref{tab:specialcases} then gives the conditions required for unlinked ($R=1/2$) neo-W invasion when there is some ancestral sex-linkage ($r<1/2$; e.g., the selected locus is on the ancestral sex chromosome and the novel sex-determining locus arises on an autosome). 
These special cases indicate that neo-W invasion occurs for a large fraction of the parameter space, even though the neo-W uncouples the sex-determining locus from a locus under selection. 
Fig \ref{fig:Combination_Centimorgans} then demonstrates that under these conditions neo-W alleles can spread when they are more loosely \textit{or} more closely linked to the locus that experiences haploid selection, i.e., Conclusions 3B and 3C (c.f., Fig \ref{fig:SexAntagTighter}A for diploid sexually-antagonistic selection alone). 
%When the neo-W is more closely linked to the haploid selected locus its invasion increases (or creates anew) sex-ratio bias (Fig \ref{fig:SexRatioBad}A).
%Under the same selection scenario (drive in males), a neo-Y would invade an ancestrally ZW system
% shows that when a neo-W invades that increases linkage with a haploid selected locus, it biases the sex-ratio.
%, while those that decrease sex-linkage equalize the sex-ratio. 
%When sex-linkage decreases ($R>r$), neo-W alleles can benefit from Y-associated alleles that have higher fitness in the female diploid phase but then dissociate from these alleles through recombination before any subsequent haploid selection.

@@

%Previous research has shown that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), alleles at a new, unlinked sex-determining locus can invade and restore equal sex ratios~\cite{Kozielska:2010vm}. 
We can also compare transitions in genetic sex-determination where sex-ratio bias increases, decreases, or remains equal. 
For example, if there is meiotic drive in males only ($\alpha^\male_\Delta \neq 0$, $\alpha^\female_\Delta=0$), without gametic competition ($t^\female=t^\male=0$) the zygotic sex ratio is initially biased only when the ancestral sex-determining system is XY (Fig \ref{fig:model_outline}A and Fig \ref{fig:SexRatioBad}A) and not ZW (Fig \ref{fig:SexRatioBad}B).
%If the ancxestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Fig \ref{fig:Combination_Turnover}D).
If Fisherian sex-ratio selection were dominant, we would thus expect a difference in the potential for XY to ZW and ZW to XY transitions. 
%However, to leading order with selection weak relative to recombination, we find that sex-ratio selection favours the spread of a neo-W by an amount equl to the spread of a neo-Y into a ZW system (relabeling XY as the heterogametic sex and reversing the labels of males and females in Table \ref{tab:specialcases})
However, invasion by a neo-W allele into an XY system and invasion by a neo-Y allele into a ZW system occur under the same conditions ($\lambda_{Y'}^{(XY)} = \lambda_{W'}^{(ZW)}$ and $\lambda_{W'}^{(XY)}=\lambda_{Y'}^{(ZW)}$, at least to order $\epsilon^2$), implying that,% surprisingly,
\vspace{0.5cm}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Tight linkage allows looser sex-linkage to evolve, even in absence of haploid selection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\centerline{\includegraphics[width=1.4\linewidth, right]{../Plots/PositionPlot_SexAntagTighter.eps}}
\caption{
{\bf Transitions between XY and ZW systems can occur even when the new sex-determining locus is less tightly linked to a locus under sexually-antagonistic selection (no haploid selection).}
In panel A, linkage is initially tight relative to selection and a neo-W can invade even when it is less tightly linked with the selected locus ($r<R$; unshaded region around \textbf{*}).
In panel B, linkage is loose enough relative to selection that the analytical results assuming weak selection hold, and a neo-W allele can only invade when it arises at a locus more tightly linked with the selected locus ($R<r$; shaded region).
In panel C we vary the recombination rate between the neo-W and the selected locus ($R$) for a fixed recombination rate between the ancestral sex-determining locus and the selected locus ($r=0.005$). 
Coloured markers show recombination rates for which the temporal dynamics of invasion are plotted in the inset (frequency of females carrying a neo-W), demonstrating that neo-W alleles can reach fixation if they are more (black) or less (red) closely linked to a locus experiencing sexually-antagonistic selection. 
A very loosely linked neo-W does not spread in this case (blue and green lines overlap and go to 0 in inset). 
Fitness parameters are: $w_{AA}^\female = 1.05$, $w_{aa}^\male = 1.2$, $w_{aa}^\female = w_{AA}^\male = 0.85$, $w_{Aa}^\circ = 1$.%,  $t^\circ = \alpha^\circ_\Delta = 0$.
%\textcolor{red}{consider removing panel A, which is repeated in Fig 3.}
%\textcolor{blue}{probably better to remove from panel 3, eh?}
}
\label{fig:SexAntagTighter}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{equation}
\lambda_{Y'}^{(XY)} = 1 + \frac{1}{4}\bar{p}(1-\bar{p}){S_{A}}^2\frac{ \left( r-R \right) }{r R}+O\left(\epsilon^3 \right) 
\label{eq:lambda_neoY}
\end{equation}

@@

%The new sex-determining allele, $m$, will spread in an XY system if $\lambda_{m}^{(XY)}>1$. 
Eq \eqref{eq:lambda_neoY} demonstrates that, under weak selection, a neo-Y allele will invade an XY system ($\lambda_{Y'}^{(XY)}>1$) if and only if it is more closely linked to the selected locus than the ancestral sex-determining locus (i.e., if $R<r$).%; note that $\bar{p}(1-\bar{p})S_{A}^2$ is strictly positive as long as $\mathbf{A}$ is polymorphic). 
This echoes our results above where a neo-Y could never invade if $r\approx0$. 
It is also consistent with the results of \cite{vanDoorn:2007eu}, who considered diploid selection only and also found that cis-GSD transitions can only occur when the new sex-determining locus is more closely linked to a locus under sexually-antagonistic selection. 
%Because tighter linkage between the sex-determining region and a selected locus evolves during these transitions, stronger associations between selected alleles and diploid sexes can build up by selection (e.g., between male-beneficial alleles and the male-determining allele, Y).%, increasing diploid mean fitness (Fig \ref{fig:Combination_MeanFit}).

@@

\noindent
Similarly, a neo-W is typically favoured when it is more closely linked to the selected locus than the ancestral sex-determining region is, ($R<r$, e.g., Figs \ref{fig:SexAntagTighter}B and \ref{fig:Combination_Centimorgans}).
Specifically, if the selected locus ($\mathbf{A}$) is ancestrally autosomal ($r=1/2$, leading to $ \hat{p}^\male_Y-\hat{p}^\male_X =0$), a neo-W linked to the selected locus ($R<1/2$) will spread ($\lambda_{W'}^{(XY)}>1$).

@@

\noindent
In the absence of haploid selection ($t^\circ=\alpha^\circ_{\Delta}=0$), Eq \eqref{eq:lambda_neoW} indicates that trans-GSD transitions can occur if and only if the new sex-determining locus is more closely linked to a locus under selection, $R<r$, as found by \cite{vanDoorn:2010hu}. 
However, with haploid selection and some ancestral sex-linkage ($r<1/2$; allowing allele frequency differences on the X and Y), the term in square brackets in Eq \eqref{eq:lambda_neoW} can be positive.
This leads to

@@

To clarify the parameter space under which neo-W alleles spread despite looser linkage with the selected locus ($R>r$), we focus on cases where dominance coefficients are equal in the two sexes, $h^\female=h^\male$, and haploid selection only occurs in one sex (e.g., during male meiosis only).
Table \ref{tab:specialcases} then gives the conditions required for unlinked ($R=1/2$) neo-W invasion when there is some ancestral sex-linkage ($r<1/2$; e.g., the selected locus is on the ancestral sex chromosome and the novel sex-determining locus arises on an autosome). 
These special cases indicate that neo-W invasion occurs for a large fraction of the parameter space, even though the neo-W uncouples the sex-determining locus from a locus under selection. 
Fig \ref{fig:Combination_Centimorgans} then demonstrates that under these conditions neo-W alleles can spread when they are more loosely \textit{or} more closely linked to the locus that experiences haploid selection, i.e., Conclusions 3B and 3C (c.f., Fig \ref{fig:SexAntagTighter}A for diploid sexually-antagonistic selection alone). 
%When the neo-W is more closely linked to the haploid selected locus its invasion increases (or creates anew) sex-ratio bias (Fig \ref{fig:SexRatioBad}A).
%Under the same selection scenario (drive in males), a neo-Y would invade an ancestrally ZW system
% shows that when a neo-W invades that increases linkage with a haploid selected locus, it biases the sex-ratio.
%, while those that decrease sex-linkage equalize the sex-ratio. 
%When sex-linkage decreases ($R>r$), neo-W alleles can benefit from Y-associated alleles that have higher fitness in the female diploid phase but then dissociate from these alleles through recombination before any subsequent haploid selection.

@@

We can also compare transitions in genetic sex-determination where sex-ratio bias increases, decreases, or remains equal. 
For example, if there is meiotic drive in males only ($\alpha^\male_\Delta \neq 0$, $\alpha^\female_\Delta=0$), without gametic competition ($t^\female=t^\male=0$) the zygotic sex ratio is initially biased only when the ancestral sex-determining system is XY (Figs \ref{fig:model_outline}B and \ref{fig:SexRatioBad}A) and not ZW (Figs \ref{fig:model_outline}B and \ref{fig:SexRatioBad}B).
%If the ancxestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Fig \ref{fig:Combination_Turnover}D).
If Fisherian sex-ratio selection were dominant, we would thus expect a difference in the potential for XY to ZW and ZW to XY transitions. 
%However, to leading order with selection weak relative to recombination, we find that sex-ratio selection favours the spread of a neo-W by an amount equl to the spread of a neo-Y into a ZW system (relabeling XY as the heterogametic sex and reversing the labels of males and females in Table \ref{tab:specialcases})
However, invasion by a neo-W allele into an XY system and invasion by a neo-Y allele into a ZW system occur under the same conditions ($\lambda_{Y'}^{(XY)} = \lambda_{W'}^{(ZW)}$ and $\lambda_{W'}^{(XY)}=\lambda_{Y'}^{(ZW)}$, at least to order $\epsilon^2$), implying that,% surprisingly,
\vspace{0.5cm}
